---
title: "Test"
author: "Steph Gervasi"
date: "5/13/2019"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    theme: united
    toc: yes
---
# Load package for this demo
```{r}
data(iris)
head(iris)
```

# Run very short EDA
```{r}
str(iris)
head(iris)
levels(iris$Species)
```

# Summary Stats
```{r}
# install.packages("tidyverse")
library(tidyverse)
group_by(iris, Species) %>%
  summarise(
    count = n(),
    mean = mean(Sepal.Length, na.rm = TRUE),
    sd = sd(Sepal.Length, na.rm = TRUE)
  )
```

# Visualizations
## Boxplot with ggpubr
```{r}
# install.packages("ggpubr")
library(ggpubr)
ggboxplot(iris, x = "Species", y = "Sepal.Length", 
          color = "Species", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("setosa", "versicolor", "virginica"),
          ylab = "Sepal Length (in)", xlab = "Flower Species Name")
```

## Boxplot with ggplot2
```{r}
library(ggplot2)
ggplot(iris, aes(x=Species, y=Sepal.Length, fill=Species)) + 
  geom_boxplot() + theme_classic() + theme(legend.position = "top") + 
  labs(x="Flower Species Name", y = "Sepal Length  (in)")
```


## Dot and line plot with ggpubr
```{r}
ggline(iris, x = "Species", y = "Sepal.Length", 
       add = c("mean_se", "jitter"), 
       order = c("setosa", "versicolor", "virginica"),
       ylab = "Sepal Length (in)", xlab = "Flower Species Name")
```

# Run group comparison analysis **(ANOVA)**
```{r}
# Compute the analysis of variance
aov1 <- aov(Sepal.Length ~ Species, data = iris)

# Summary of the analysis
summary(aov1)

# Posthoc tests
TukeyHSD(aov1)
```

# Check assumptions of analysis/model
## Check homogeneity of variance assumption
```{r}
plot(aov1, 1)
#plot(aov1)

library(car)
leveneTest(Sepal.Length ~ Species, data = iris)
# From the output we can see that the p-value is less than the significance level of 0.05. This means that there is evidence to suggest that the variance across groups is statistically significantly different. Therefore, we cannot assume the homogeneity of variances in the different treatment groups.

# It is possible to run a Welch's ANOVA where the assumption of homogeneity of variance is relaxed. It is implemented as follows in R:

oneway.test(Sepal.Length ~ Species, data = iris)

# Now pairwise t tests on this output:

pairwise.t.test(iris$Sepal.Length, iris$Species,
                 p.adjust.method = "BH", pool.sd = FALSE)
```


## Check normality assumption
```{r}
plot(aov1, 2)

# Additional check for normality besides QQ plot produced above.

# Extract the residuals
aov_residuals <- residuals(object = aov1)

# Run Shapiro-Wilk test
shapiro.test(x = aov_residuals)

# If this had been significant, it would indicate that we were in violation of the assumption of normality. We could transform our response variable, or could proceed with another, non-parametric, test. A non-parametric alternative to one-way ANOVA is Kruskal-Wallis rank sum test, which can be used when ANOVA assumptions are not met.

kruskal.test(Sepal.Length ~ Species, data = iris)

```

# Some exploration with plotting in ggpubr
## Multiple groups and Faceting
```{r}
data(ToothGrowth)
head(ToothGrowth)
p <- ggboxplot(ToothGrowth, x = "dose", y = "len",
      color = "supp", palette = c("#00AFBB", "#E7B800"),
      ylab = "Length", xlab = "Dose")
print(p)

facet(p, facet.by = "supp")
```

## Density plots
```{r}
# Density plot with mean lines and marginal rug
# :::::::::::::::::::::::::::::::::::::::::::::::::::
# Change outline and fill colors by groups
# Use custom palette
ggdensity(iris, x = "Sepal.Length",
   add = "mean", rug = TRUE,
   color = "Species", fill = "Species",
   palette = c("#00AFBB", "#E7B800", "#FC4E07"))
```

## Histograms
```{r}
# Histogram plot with mean lines and marginal rug
# :::::::::::::::::::::::::::::::::::::::::::::::::::
# Change outline and fill colors by groups ("sex")
# Use custom color palette
gghistogram(iris, x = "Sepal.Length",
   add = "mean", rug = TRUE,
   color = "Species", fill = "Species",
   palette = c("#00AFBB", "#E7B800", "#FC4E07"))
```

## Boxplots with jittered points
```{r}
# Box plots with jittered points
# :::::::::::::::::::::::::::::::::::::::::::::::::::
# Change outline colors by groups: dose
# Use custom color palette
# Add jitter points and change the shape by groups
 p <- ggboxplot(ToothGrowth, x = "dose", y = "len",
                color = "dose", palette =c("#00AFBB", "#E7B800", "#FC4E07"),
                add = "jitter", shape = "dose")
 p
```

## Kruskal Wallis
```{r}
head(ToothGrowth)
str(ToothGrowth)

# Convert the dose variable to a factor
tg <- ToothGrowth
tg$dose <- as.factor(tg$dose)
head(tg)
str(tg)

# Perform the test
kruskal.test(len ~ dose, data = tg)

# multiple comparisons with Dunn test
#install.packages("FSA")
library(FSA)
dunnTest(len ~ dose,
              data=tg,
              method="bh")    # Can adjust p-values;
                              # See ?p.adjust for options

# multiple comparisons using wilcoxon test :THIS APPEARS TO BE WHAT IS USED IN THE GRAPHS BELOW!!!
pairwise.wilcox.test(tg$len,
                          tg$dose,
                          p.adjust.method="none")
                              # Can adjust p-values;
                              # See ?p.adjust for options

```


## Boxplots with stats output!
```{r}
# Add p-values comparing groups
 # Specify the comparisons you want
my_comparisons <- list( c("0.5", "1"), c("1", "2"), c("0.5", "2") )
p + stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 50)                   # Add global p-value
```

## Violin with box plots including stats output
```{r}
# Violin plots with box plots inside
# :::::::::::::::::::::::::::::::::::::::::::::::::::
# Change fill color by groups: dose
# add boxplot with white fill color
ggviolin(ToothGrowth, x = "dose", y = "len", fill = "dose",
         palette = c("#00AFBB", "#E7B800", "#FC4E07"),
         add = "boxplot", add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons, label = "p.signif")+ # Add significance levels
  stat_compare_means(label.y = 50)                                      # Add global the p-value 
```

## Bar plots and references
```{r}
# Load data
data("mtcars")
dfm <- mtcars
# Convert the cyl variable to a factor
dfm$cyl <- as.factor(dfm$cyl)
# Add the name colums
dfm$name <- rownames(dfm)
# Inspect the data
head(dfm[, c("name", "wt", "mpg", "cyl")])

ggbarplot(dfm, x = "name", y = "mpg",
          fill = "cyl",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "desc",          # Sort the value in dscending order
          sort.by.groups = FALSE,     # Don't sort inside each group
          x.text.angle = 90           # Rotate vertically x axis texts
          )


# sort bars in each group

ggbarplot(dfm, x = "name", y = "mpg",
          fill = "cyl",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",           # Sort the value in dscending order
          sort.by.groups = TRUE,      # Sort inside each group
          x.text.angle = 90           # Rotate vertically x axis texts
          )

# See these references:
# http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/
# http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization
# http://www.sthda.com/english/wiki/one-way-anova-test-in-r
# https://rcompanion.org/rcompanion/d_06.html
  
```

